---
title: "SFB Blind Bid Waivers"
output: html_document
editor_options: 
  chunk_output_type: inline
---

## Correlation of early season blind bidding prices to future fantasy points/ranks - is blowing all your money early a good thing?

Are we good at spending our waiver wire money?

We = ScottFishBowl participants, as a proxy for the fantasy football industry

------------------------------------------------------------------------

Base question: average bid per player across the whole season - relationship with PPG/rank

Time factor - week one/month-one bids correlate with PPG/rank

Identify user patterns - are there people who churn more, or people who spend all their money, or people who save up and spend later?

Position-related patterns - QB/RB/WR/TE effects/correlations

Year-over-year patterns - requires namecleaning

Fans - comparison of expert vs "Fan -"

Split franchises into top-third/middle-third/bottom-third of overall standings performance

Look at percentage of blind bid spent

```{r setup}

suppressPackageStartupMessages({
  
  options(ffscrapr.cache = "filesystem")
  options(dplyr.summarise.inform = FALSE)
  
  library(tidyverse)
  library(arrow)
  library(ffscrapr)
  library(janitor)
  library(glue)
  library(lubridate)
  library(hrbrthemes)
  library(GGally)
  
  import_roboto_condensed()
  
})

```

Load data from repo

```{r}


setwd(here::here())

transactions <- open_dataset("data/transactions") %>% collect()

standings <- open_dataset("data/standings") %>% collect()

player_sh <- open_dataset("data/league") %>% collect() %>% 
  filter(sfb_type == "official") %>% 
  distinct(year, .keep_all = TRUE) %>%
  select(year, league_id) %>% 
  mutate(
    scoring = map2(year,
                   league_id, 
                   ~ mfl_connect(.x,
                                 .y, 
                                 user_agent = paste0("dynastyprocess/",.y),
                                 rate_limit_number = 2,
                                 rate_limit_seconds = 3) %>% 
                     ff_scoringhistory(season = .x))
  ) %>% 
  select(-league_id) %>% 
  unnest(scoring) %>% 
  left_join(dp_playerids() %>% select(gsis_id,mfl_id2 = mfl_id),
            by = "gsis_id") %>% 
  group_by(season, gsis_id,mfl_id,mfl_id2) %>% 
  count() %>% 
  ungroup() %>% 
  transmute(mfl_id = coalesce(mfl_id2,mfl_id),
            season,
            n)

player_scores <- open_dataset("data/league") %>% collect() %>% 
  filter(sfb_type == "official") %>% 
  distinct(year, .keep_all = TRUE) %>%
  select(year, league_id) %>% 
  mutate(
    scoring = map2(year,
                   league_id, 
                   ~ mfl_connect(.x,
                                 .y, 
                                 user_agent = paste0("dynastyprocess/",.y),
                                 rate_limit_number = 2,
                                 rate_limit_seconds = 3) %>% 
                     ff_playerscores(season = .x, week = "AVG"))
  ) %>% 
  select(-league_id) %>% 
  unnest(scoring) %>% 
  left_join(player_sh, by = c("player_id"="mfl_id","season")) %>% 
  filter(!is.na(n), n>=4) %>% 
  group_by(season, pos) %>% 
  mutate(pos_rank = rank(desc(points),ties.method = "min")) %>% 
  ungroup() %>% 
  select(season, player_id, player_name, pos, team, points, pos_rank)


bbid_waivers <- transactions %>% 
  filter(sfb_type == "official",
         type_desc == "added") %>% 
  select(season = year,
         timestamp,
         league_id,
         league_name,
         franchise_id,
         franchise_name,
         bbid_spent,
         player_id,
         player_name,
         pos,
         team
  ) %>% 
  mutate(bbid_spent = replace_na(bbid_spent,0)) %>% 
  left_join(
    player_scores %>% select(season, player_id, points, pos_rank),
    by = c("season", "player_id")
  )

```

Plot raw bbid_spent against the PPG and against pos rank

```{r}

bbid_waivers %>% 
  mutate(season = as.character(season)) %>% 
  group_by(season, player_id, points, pos_rank) %>% 
  summarise(bbid_spent = mean(bbid_spent, na.rm = TRUE)) %>% 
  ggplot(aes(x = bbid_spent, y = points, group = season, colour = season)) +
  geom_point(alpha = 0.2) +
  geom_smooth(se = FALSE, method = "lm") + 
  scale_x_log10() +
  labs(
    title = "Average BBID Spent vs PPG",
    caption = "ScottFishBowl, 2016-2020"
  ) +
  theme_modern_rc()

bbid_waivers %>% 
  mutate(season = as.character(season)) %>% 
  group_by(season, player_id, points, pos_rank) %>% 
  summarise(bbid_spent = mean(bbid_spent, na.rm = TRUE)) %>% 
  ggplot(aes(x = bbid_spent, y = pos_rank, group = season, colour = season)) +
  geom_point(alpha = 0.2) +
  geom_smooth(se = FALSE, method = "lm") + 
  scale_x_log10() +
  scale_y_reverse() +
  labs(title = "Average BBID Spent vs Position Rank",
       caption = "ScottFishBowl, 2016-2020") +
  theme_modern_rc()

bbid_waivers %>% 
  mutate(season = as.character(season)) %>% 
  group_by(season, player_id, pos, points, pos_rank) %>% 
  summarise(bbid_spent = mean(bbid_spent, na.rm = TRUE)) %>% 
  ggplot(aes(x = bbid_spent, y = pos_rank, group = pos, colour = pos)) +
  geom_point(alpha = 0.2) +
  geom_smooth(se = FALSE, method = "lm") + 
  scale_x_log10() +
  scale_y_reverse() +
  coord_cartesian(ylim = c(100,0)) +
  # ylim(c(32,0)) +
  labs(title = "Average BBID Spent vs Position Rank",
       subtitle = "by Position",
       caption = "ScottFishBowl, 2016-2020") +
  theme_modern_rc()

bbid_waivers %>% 
  mutate(season = as.factor(season)) %>% 
  group_by(season, player_id, player_name, pos, team, points, pos_rank) %>% 
  summarise(bbid_spent = mean(bbid_spent, na.rm = TRUE)) %>%
  ungroup() %>% 
  mutate(log_bbidspent = log(bbid_spent+1, base = 10)) %>%
  # mutate(points = replace_na(points,0),
  # pos_rank = replace_na(pos_rank, 0)) %>% 
  # group_by(season, pos) %>% 
  # mutate_at(c("bbid_spent","log_bbidspent","points","pos_rank"), ~ (.x -min(.x))/(max(.x)-min(.x))) %>% 
  # ungroup() %>% 
  select(season, player_id,player_name, pos, team, log_bbidspent, bbid_spent, points, pos_rank) %>%
  ggpairs(columns = 6:9, ggplot2::aes(colour=pos))

```

How well does spending correlate with points and PPG

```{r}

bbid_waivers %>% 
  mutate(season = as.character(season)) %>%
  filter(!is.na(pos)) %>%
  group_by(season, player_id, points, pos_rank, pos) %>% 
  summarise(bbid_spent = mean(bbid_spent)) %>% 
  ungroup() %>% 
  mutate(log_bbidspent = log(bbid_spent+1, base = 10)) %>%
  group_by(pos) %>% 
  summarise(
    corr_points = cor(bbid_spent, points, use = "complete.obs"),
    corr_posrank = cor(bbid_spent, pos_rank, use = "complete.obs"),
    corr_points_log10 = cor(log_bbidspent, points, use = "complete.obs"),
    corr_posrank_log10 = cor(log_bbidspent, pos_rank, use = "complete.obs")
  )

```
